DRAFT: Week 3: Degree and Centrality

### 2022-03-05 docs/3. SNA - Assignments/sna_Week3/DRAFT-week3-Assignment.Rmd
3.5.22

## Smuggler's Index: Network Attributes 
```{r Network Size}

@@

@@

@@

list(
  triad_census(igEL)
)

```
plot(
  triad_census(igEL)
)

@@

#     total number of all four triad types returned by triad census
sum(
  sna::triad.census(
    snEL,
    mode="graph"))

#     Classify all Triad in the network: Undirected Network
 sna::triad.census(
     snEL,
     mode="graph")

@@

sna::triad.classify(
     snEL,
     tri=c(1,2,3),
     mode==c("graph")
   )

@@

 summary(  
  transitivity(
   igEL,
   type = c("undirected"),
   vids = NULL, #  all vertices are considered.
   weights = NULL,
   isolates = c("NaN", "zero"))
)

summary( 
 transitivity(igEL,
             type = c("globalundirected"),
             vids = NULL, #  all vertices are considered.
             weights = NULL,
             isolates = c("NaN", "zero"))
)

summary(
  transitivity(igEL,
             type = c("localundirected"),
             vids = NULL, #  all vertices are considered.
             weights = NULL,
              isolates = c("NaN", "zero"))
)
#   c("undirected", "global", "globalundirected", "localundirected", "local","average", "localaverage","localaverageundirected", "barrat", "weighted")

```
plot(
  transitivity(igEL,
             type = c("localundirected"),
             vids = NULL, #  all vertices are considered.
             weights = NULL,
              isolates = c("NaN", "zero"))
)
```{r statnet: Transitivity:, echo=FALSE}
library(statnet)
library(sna)

@@

###  Statnet:   Path Length and Geodesic Distance

@@

 gd_snEL <- geodist(
    snEL, 
    inf.replace=.5, #geodesics do not exist, substitute value
    count.paths=TRUE, 
    predecessors=T, # return predecessor list
    ignore.eval=F, # ignore edge values
    na.omit=F) # remove NA values

summary(gd_snEL)
view(
  list(gd_snEL))

  #Examine the number of geodesics
snEL_gdCounts <- gd_snEL$counts

@@

```
gplot(gd_snEL)
gplot(snEL_gdist)

@@

summary(
  reachability(
    snEL,
    geodist.precomp=NULL,
    return.as.edgelist=FALSE,
    na.omit=FALSE)
)
```

```{r STATNET:Path: Geodesic number & lengths}
library(statnet)
library(sna)

summary(geodist(snEL, 
        inf.replace=Inf, 
        count.paths=TRUE, # include geodesics 
        predecessors=FALSE,# include predecessor list
        ignore.eval=FALSE, # ignore edge values
        na.omit=FALSE) # remove NA-values edges
        )

#Examine the number of geodesics
gd_snEL$counts

#Examine the geodesic distances
gd_snEL$gdist

  summary(gd_snEL)
  gplot(gd_snEL)

summary(kpath.census(snEL, 
             maxlen = 5, 
             mode = "graph",
             tabulate.by.vertex = TRUE, 
             path.comembership = c("none"), # c("none", "sum","bylength"), 
             dyadic.tabulation = c("none") # c("none", "sum", "bylength"))
))
              ##**"sum"* returns a vertex by vertex matrix of source/destination path counts
             ##**"bylength"* disaggregates these counts by path length. 
             ##**"none"* disables this computation.

```
- Aggregate counts of paths/cycles by length (i.e., k).

@@

###  IGRAPH:   Path Length and Geodesic Distance

```{r iGraph:Nodes: Path: Length, echo=TRUE}
library(igraph)

@@

  The shortest path to describe the overall network structure 
```{r iGraph:Network: Path: Shortest, echo=TRUE}

#     find average shortest path for network
    average.path.length(igEL,directed=F)

@@

```{r iGraph:Nodes: Path: Length, echo=TRUE}

library(igraph)
# Head
list(
  head(V(igEL)$name,10),
  tail(V(igEL)$name,10)
  )

#     Calculate distance between two nodes using unweighted edges

    distances(igEL,"AKU AKU" ,  "SINGAPORE SLING",weights=NA)
    distances(igEL,"AKU AKU" ,  "HONDO HATTIE'S JUNGLE PUNCH!",weights=NA)
    distances(igEL,"AKU AKU" ,"KAHIKO PUNCH" ,weights=NA)
    distances(igEL,"AKU AKU" , "PASSION FRUIT HONEY"   ,weights=NA)
    distances(igEL,"AKU AKU" , "TRIUMVIRATE PUNCH"   ,weights=NA)
    distances(igEL,"AKU AKU" ,"EUREKA PUNCH" ,weights=NA)
    distances(igEL,"AKU AKU" , "YULETIDAL WAVE" ,weights=NA)
    distances(igEL,"AKU AKU" ,"RUMBUSTION PUNCH"   ,weights=NA)
    distances(igEL,"AKU AKU" ,"PUKA PUNCH",weights=NA)

@@

```{r iGraph:Nodes: Path: Shortest, echo=TRUE}

#     isolating the distances between specific node

## Note: manually tell igraph to ignore edge weights.

@@

aku_PFhoney <- all_shortest_paths(igEL,
                     "AKU AKU", 
                     "PASSION FRUIT HONEY",
                     weights=NA)$res

@@

aku_Puka <-   all_shortest_paths(igEL,
                     "AKU AKU" ,
                     "PUKA PUNCH",
                     weights=NA)$res
```

### Statnet: Component STructure, Isolates, Density
```{r Statnet: Isolates, echo=TRUE}

              ## Isolates:

@@

 # Is Ego an Isolate?
 is.isolate(snEL, ego_exp, g=1, diag=FALSE)

 # Adds n isolates to the graph (or graphs)
 add.isolates(snEL, n_umber, return.as.edgelist = FALSE)

@@

@@

### Igraph: Component STructure, Isolates,
```{r iGraph: Undirected: Component Structure, echo=TRUE}
                ## Component Structure:

    # Number of components
 igraph::components(igEL)$no 
      ##**NOTE:** $no prevents a list of components from being printed

# Size of each component
 igraph::components(igEL)$csize

    #get network density **with** loops
 network.density(igEL)
 class(igEL)
          ##**NOTE** provides an incorrect density calculation 
          ## as the default assumes that loops are not included but does not remove them

```
 **Igraph** provides an incorrect density calculation as the default assumes  that loops are not included but does not remove them.

 Therefore, the default calculation includes the loops in the numerator but not the denominator of the density calculation.

>>>>

## Smuggler's Index: Network Attributes

```{r Network Size}

@@

```
plot.network(snEL)
igraph::plot.igraph(igEL)

@@

    #get network density **with** loops
 igraph::graph.density(igEL)
 edge_density(igEL)
          ##**NOTE** provides an incorrect density calculation 
          ## as the default assumes that loops are not included but does not remove them

@@

igEL_triList <- 
  triad_census(igEL)

```
plot(igEL_triList)

@@

#     total number of all four triad types returned by triad census

#     Classify all Triad in the network

  sna::triad.census(
    snEL,
    mode="graph")

view(
 sna::triad.census(
     snEL,
     mode="digraph")
))

@@

sna::triad.classify(
     snEL,
     tri=c(1,2,3),
     mode==c("graph"))

sna::triad.classify(
     snEL,
     tri=c(1:16),
     mode==c("digraph"))

```
plot(view(
 sna::triad.census(
     snEL,
     mode="digraph")
))

plot(view(
sna::triad.census(snEL,mode="graph")))

@@

 summary(  
  transitivity(
   igEL,
   type = c("undirected"),
   vids = NULL, #  all vertices are considered.
   weights = NULL,
   isolates = c("zero"))
)

summary( 
 transitivity(igEL,
             type = c("globalundirected"),
             vids = NULL, #  all vertices are considered.
             weights = NULL,
             isolates = c("zero"))
)

summary(
  transitivity(igEL,
             type = c("localundirected"),
             vids = NULL, #  all vertices are considered.
             weights = NULL,
              isolates = c("zero"))
)
#   c("undirected", "global", "globalundirected", "localundirected", "local","average", "localaverage","localaverageundirected", "barrat", "weighted")

```{r statnet: PLOT: Transitivity:, echo=FALSE}
library(statnet)
library(sna)

@@

###  Statnet:   Path Length and Geodesic Distance: Networks

@@

View(
geodist(snEL))

 gd_snEL <- geodist(
    snEL, 
    inf.replace=.25, #geodesics do not exist, substitute value
    count.paths=TRUE, 
    predecessors=F, # return predecessor list
    ignore.eval=T, # ignore edge values
    na.omit=T) # remove NA values

  #Examine the number of geodesics
snEL_gdCounts <- gd_snEL$counts
  sum(snEL_gdCounts)
  log_mean_exp(snEL_gdCounts)
  log_sum_exp(snEL_gdCounts)

@@

  sum(snEL_gdist)
  log_mean_exp(snEL_gdist)
  log_sum_exp(snEL_gdist)

```
gplot(gd_snEL) in 
gplot(snEL_gdist)

@@

summary(
  reachability(
    snEL,
    geodist.precomp= gd_snEL,
    return.as.edgelist=F,
    na.omit=F)
)
```

```{r STATNET:PLOT: Reachability}
library(statnet)
library(sna)

gplot(
  reachability(
    snEL,
    geodist.precomp= gd_snEL,
    return.as.edgelist=F,
    na.omit=F)
)
```

    K-Path Census
- Aggregate counts of paths/cycles by length (i.e., k).

@@

summary(kpath.census(snEL, 
             maxlen = 5, 
             mode = "graph",
             tabulate.by.vertex = T, 
             path.comembership = c("sum"), # c("none", "sum","bylength"),
             dyadic.tabulation = c("sum") # c("none", "sum", "bylength"))
))
              ##**"sum"* returns a vertex by vertex matrix of source/destination path counts
             ##**"bylength"* disaggregates these counts by path length. 
             ##**"none"* disables this computation.
```

```{r STATNET: PLOT: k Path Census}
library(statnet)
library(sna)

gplot(kpath.census(snEL, 
             maxlen = 10, 
             mode = "graph",
             tabulate.by.vertex = T, 
             path.comembership = c("bylength"), # c("none", "sum","bylength"),
             dyadic.tabulation = c("sum") # c("none", "sum", "bylength"))
))
              ##**"sum"* returns a vertex by vertex matrix of source/destination path counts
             ##**"bylength"* disaggregates these counts by path length. 
             ##**"none"* disables this computation.

@@

###  IGRAPH:    Path Length and Geodesic Distance: Networks & Nodes

  The shortest path to describe the overall network structure 
```{r iGraph: Path: Network & Nodes, echo=TRUE}
library(igraph)

#     find average shortest path for network
    average.path.length(igEL,
                        directed=F,
                       ig_histPath)

    igraph::path.length.hist(igEL)$no

#  shortest_paths calculates one shortest path (the path itself, and not just its length) from or to the given vertex.

  shortest_paths(
  igEL,
  from=1,
  to = V(igEL),
  mode = c("all"),
  weights = NULL,
  output = c("epath"), #c("vpath", "epath", "both")
  predecessors = FALSE,
  inbound.edges = FALSE)$no

    all_shortest_paths(
      igEL,
      from = 1, # Vertex ID: AkuAku
      to = V(igEL),
      mode = c("all"),
      weights = NULL)$no

##**From=(vertex id)*	
    ##Numeric constant, the vertex from or to the shortest paths

##**mode=c("out", "all", "in")*

##**output= *	
    ## defines how to report the shortest paths. 
  #**“vpath”* the vertices along the paths are reported, this form was used prior to igraph version 0.6. 
  #**“epath”* edges along the paths are reported. 
  #*“both”* both forms are returned, in a named list with components “vpath” and “epath”.

@@

```{r iGraph: PLOT: PATH: Network & Nodes, echo=TRUE}
library(igraph)

 ig_histPath <-
view(
    igraph::path.length.hist(igEL)
)
#  shortest_paths calculates one shortest path (the path itself, and not just its length) from or to the given vertex.

ig_ShortPath <-
  shortest_paths(
  igEL,
  from=1,
  to = V(igEL),
  mode = c("all"),
  weights = NULL,
  output = c("epath"), #c("vpath", "epath", "both")
  predecessors = FALSE,
  inbound.edges = FALSE)

ig_AllShort <- 
    all_shortest_paths(
      igEL,
      from = 1, # Vertex ID: AkuAku
      to = V(igEL),
      mode = c("all"),
      weights = NULL)$res

##**From=(vertex id)*	
    ##Numeric constant, the vertex from or to the shortest paths

##**mode=c("out", "all", "in")*

##**output= *	
    ## defines how to report the shortest paths. 
  #**“vpath”* the vertices along the paths are reported, this form was used prior to igraph version 0.6. 
  #**“epath”* edges along the paths are reported. 
  #*“both”* both forms are returned, in a named list with components “vpath” and “epath”.

 class(ig_histPath)
  class(ig_ShortPath)
  class(ig_AllShort)
```   
  Distance between two or more Nodes
```{r iGraph: Distances: Network & Nodes, echo=TRUE}

library(igraph)
# distances calculates the length of all the shortest paths from or to the vertices in the network.

 mean_distance(igEL, directed = F, unconnected = TRUE)

view(
  distance_table(igEL, directed = F)
)

view(
  igraph::distances(igEL)
)

### distances between specific nodes

    distances(igEL,"AKU AKU" ,  "SINGAPORE SLING",weights=NA)
    distances(igEL,"AKU AKU" ,  "HONDO HATTIE'S JUNGLE PUNCH!",weights=NA)
    distances(igEL,"AKU AKU" ,"KAHIKO PUNCH" ,weights=NA)
    distances(igEL,"AKU AKU" , "JET PILOT"   ,weights=NA)
    distances(igEL,"AKU AKU" , "TRIUMVIRATE PUNCH"   ,weights=NA)
    distances(igEL,"AKU AKU" ,"EUREKA PUNCH" ,weights=NA)
    distances(igEL,"AKU AKU" , "YULETIDAL WAVE" ,weights=NA)
    distances(igEL,"AKU AKU" ,"RUMBUSTION PUNCH"   ,weights=NA)
    distances(igEL,"AKU AKU" ,"PUKA PUNCH",weights=NA)

```{r iGraph:Nodes: GRAPH:  Length between Nodes, echo=TRUE}
library(igraph)

class(aku_Puka)

View(
  list(
    aku_Sing,
    aku_Hondo,
    aku_Kahiko,
    aku_JetPilot,
    aku_Trium,
    aku_Eureka,
    aku_YuleTide,
    aku_Rumbustion,
    aku_Puka)
  )

@@

```{r iGraph:Nodes & Network: Path: Shortest , echo=TRUE}

# shortest_paths calculates one shortest path (the path itself, and not just its length) from or to the given vertex.
view(
  igraph::shortest.paths(igEL)
  )

#     isolating the distances between specific node
## Note: manually tell igraph to ignore edge weights.

@@

aku_JetPilot <- all_shortest_paths(igEL,
                     "AKU AKU", 
                     "JET PILOT",
                     weights=NA)$res

@@

aku_Puka <-   all_shortest_paths(igEL,
                     "AKU AKU" ,
                     "PUKA PUNCH",
                     weights=NA)$res

View(list(
  aku_Sing,
  aku_Hondo,
  aku_Kahiko,
  aku_JetPilot,
  aku_Trium,
  aku_Eureka,
  aku_YuleTide,
  aku_Rumbustion,
  aku_Puka)
)

```{r iGraph:Nodes: GRAPH: Shortest between Nodes, echo=TRUE}
library(igraph)

class(aku_Puka)

View(
  list(
    aku_Sing,
    aku_Hondo,
    aku_Kahiko,
    aku_JetPilot,
    aku_Trium,
    aku_Eureka,
    aku_YuleTide,
    aku_Rumbustion,
    aku_Puka)
  )

### Igraph: Component Structure, Isolates,

```{r iGraph: network density, echo=TRUE}

    #get network density **with** loops
 igraph::graph.density(igEL)
 edge_density(igEL)
          ##**NOTE** provides an incorrect density calculation 
          ## as the default assumes that loops are not included but does not remove them
```

```{r iGraph: Component Structure & Isolates, echo=TRUE}

## Component Structure:

# Number of Components in a Graph
  count_components(igEL)

# Connected components ?
 igraph::is_connected(igEL)

# Calculate the maximal (weakly or strongly) connected components of a graph

 component_distribution(igEL, 
                       cumulative =F, # relative frequency) is calculated
                       mul.size = F) # multiplied by cluster sizes
                       #**mode= c("strong"/"weak"))*

# a list with: membership by node, size of each component, and the total number of components.

 igraph::components(igEL)$no # mode = c("weak", "strong"))
      ##**NOTE:** $no prevents a list of components from being printed
summary(
  igraph::components(igEL,
                     mode = c("weak"))
)

summary(
  igraph::components(igEL,
                     mode = c("strong"))
)
 # Size of each component

 igraph::components(igEL)$csize

 # Membership of each component 
summary(  
igraph::components(igEL)$membership
)

components are all the same size, so no interesting plots

### Statnet: Isolates & Component Structure
```{r Statnet: Isolates, echo=TRUE}
library(statnet)
library(sna)
              ## Isolates:

@@

 # Is Ego an Isolate?
 is.isolate(snEL,
            1,# vertex index number (or a vector of vertices)
            g=1, 
            diag=FALSE) #diagonals (i.e., loops) contain meaningful data

 # Adds n isolates to the graph (or graphs)
 ##**add.isolates(snEL, n_umber, return.as.edgelist = FALSE)*

 #subset vertex.names attribute to get names of isolates
  ##**as.vector(SOURCE.stat%v%'vertex.names')[c(isolates(SOURCE.stat))]*

@@

```{r Statnet: Component Structure: Component Size Distribution of a Graph, echo=TRUE}
library(statnet)
library(sna)

# Number of Compnents
sna::components(snEL)

# returns a list of vectors: the number of components of graph according to size, length, and membership
summary(
component.dist(snEL)
)

summary(
component.dist(snEL,
               connected=c("unilateral"))
)

summary(
component.dist(
 snEL,
  connected=c("unilateral"))$csize
)

summary(
component.dist(
 snEL,
  connected=c("unilateral"))$cdist 
)

# identifies the component(s) of maximum order within graph
summary(
component.largest(snEL)
)

summary(
component.largest(
 snEL, 
  connected=c("weak"),
  result = c("membership"), # "membership", "graph")
  return.as.edgelist = T)
)

#  computes the component structure of the input network, and returns a vector

summary(
component.size.byvertex(snEL)
)

summary(
component.size.byvertex(
 snEL, 
  connected = c("unilateral")) # "strong", "weak", "unilateral", "recursive"
)

@@
